From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-22 paper.tex
Refining presentation of result tables

    \begin{tabular}{p{5.5cm}cccccccc}
        \toprule
        {\bf Subject} & 1 & 2 & 3 & 4 & 5 & 6 & {\bf Avg.} & {\it SD} \\
        \midrule
        {\bf Head pose tracking} (\%) & 88.2 & 83.5 & 90.5 & 83.1 & 87.9 & 85.0 & {\bf 86.4} & {\it 3.0} \\ 
        \midrule
        {\bf Ground-truth/robot match} (\%) & 61.1 & 85.5 & 80.8 & 62.2 & 63.4 & 83.9 & {\bf 72.8} & {\it 11.7}\\
        {\bf Cronbach's $\alpha$} & 0.63 & 0.81 & 0.82 & 0.59 & 0.64 & 0.78 & {\bf0.71} & {\it0.10}\\
        {\bf Cosine affinity} & 0.73 & 0.82 & 0.87 & 0.73 & 0.77 & 0.89 & {\bf 0.8} & {\it 0.07} \\
        \bottomrule
    \end{tabular}
    \label{tab:results}
\end{table*}

@@

Algorithm~\ref{alg:with-me-ness} provides a formal way of computing the level of
with-me-ness $\mathcal{W}$ between two time points $[t_{start}, t_{end}]$. A
notable difference with the original definition by Sharma~\etal is that, at a
given time $t$, the task $task(t)$ performed by the robot may elicit more than
one attentional target; thus, at a given time, more than one location can be
regarded as possible \emph{expected} focuses of attention for the human. For
example, a robot which is writing, could typically elicit gazes to its hand as
well as to its head. A human looking at either of these locations would be
considered to be \emph{with} the robot in terms of interaction. Also notable, we
exclude from the computation of $\mathcal{W}$ all of the periods of time where
the user's focus of attention can not be estimated (Algorithm~\ref{alg:with-me-ness},
line~\ref{algline:skiplosttrack}, third condition), typically because the user's
face is not visible at those times.

\fixme{mention somewhere probabilistic extensions of withmeness}

@@

During the experiment presented above (section~\ref{sec:expe}), the robot
controller associated a (set of) expected attentional targets to the task being
performed. For instance, while the robot was waiting for the child's handwriting
demonstration, the expected attentional target was the tablet (since the child
was supposed to write there), and while the robot was writing itself, the
attentional targets we expected to be either on the tablet or on the robot.
These expected targets, forming together the robot's attentional {\it a priori},
are depicted on Figure~\ref{fig:with-me-ness} as the green lines.

@@

    \begin{tabular}{p{1cm}cccccccc}
        \toprule
        Subject & 1 & 2 & 3 & 4 & 5 & 6 & Avg. & SD \\
        \midrule
        $\mathcal{W}_{g.truth}$ & 92.6 & 88.8  & 95.8 & 93.1 & 96.5 & 89.6 & 92.7 & 3.1 \\ 
        $\mathcal{W}_{robot}$ & 89.6 & 84.4 & 86.6 & 72.8 & 81.9 & 79.7 & 82.5 & 5.4 \\
        \bottomrule
    \end{tabular}
    \label{tab:results-with-me-ness}
\end{table}

According to figure \ref{fig:with-me-ness} several findings can be analyzed in
detail. The matching curve (cyan color) shows the overlapping between the
expected and the real focuses of attention. As can be noticed there is a
fluctuation at the beginning of the interaction as well as just before the
activity switching. These chaotic gaze manifestation is what we call an
automatic discovery of events, when the user experiences an uncertainty in the
present situation.

>>>>

    \begin{tabular}{p{5.5cm}cccccccc}
        \toprule
        {\bf Subject} & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
        \midrule
        {\bf Head pose tracking} (\%) & 88.2 & 83.5 & 90.5 & 83.1 & 87.9 & 85.0 & {\bf 86.4} & {\it 3.0} \\ 
        \midrule
        {\bf Ground-truth/robot match} (\%) & 61.1 & 85.5 & 80.8 & 62.2 & 63.4 & 83.9 & {\bf 72.8} & {\it 11.7}\\
        {\bf Cronbach's $\alpha$} & 0.63 & 0.81 & 0.82 & 0.59 & 0.64 & 0.78 & {\bf0.71} & {\it0.10}\\
        {\bf Cosine affinity} & 0.73 & 0.82 & 0.87 & 0.73 & 0.77 & 0.89 & {\bf 0.8} & {\it 0.07} \\
        \bottomrule
    \end{tabular}
    \label{tab:results}
\end{table*}

@@

Algorithm~\ref{alg:with-me-ness} provides a formal way of computing the level of
with-me-ness $\mathcal{W}$ between two time points $[t_{start}, t_{end}]$. A
notable difference with the original definition by Sharma~\etal is that, at a
given time $t$, the task $task(t)$ performed by the robot may elicit more than
one attentional target; thus, at a given time, more than one location can be
regarded as possible \emph{expected} focuses of attention for the human. For
example, a robot which is writing, could typically elicit gazes to its hand as
well as to its head. A human looking at either of these locations would be
considered to be \emph{with} the robot in terms of interaction\footnote{As
future work, this could easily be extended into a probabilistic model of
attention expectations.}. Also notable, we exclude from the computation of
$\mathcal{W}$ all of the periods of time where the user's focus of attention can
not be estimated (Algorithm~\ref{alg:with-me-ness},
line~\ref{algline:skiplosttrack}, third condition), typically because the user's
face is not visible at those times.

@@

During the experiment presented above (section~\ref{sec:expe}), the robot
controller associated a (set of) expected attentional targets to the task being
performed. For instance, while the robot was waiting for the child's handwriting
demonstration, the expected attentional target was the tablet (since the child
was supposed to write there), and while the robot was writing itself, the
attentional targets we expected to be either on the tablet or on the robot.
These expected targets, forming altogether the robot's attentional {\it a
priori}, are depicted on Figure~\ref{fig:with-me-ness} as the green lines. They
are used to compute the with-me-ness. With-me-ness can be calculated over the
whole interaction (Table~\ref{tab:results-with-me-ness}) or over shorter time
windows (bottom graph on Figure~\ref{fig:with-me-ness}). Shorter time windows
are interesting for two purposes: to analyse the level of with-me-ness in
relation to specific interaction episodes; to allow a measurement of
with-me-ness by the robot \emph{over the course} of the interaction
(\emph{in-the-moment} measurement). In this later case, one may typically want
to consider a sliding time window.

@@

    \begin{tabular}{p{1cm}cccccccc}
        \toprule
        Subject & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
        \midrule
        $\mathcal{W}_{g.truth}$ & 92.6 & 88.8  & 95.8 & 93.1 & 96.5 & 89.6 & {\bf 92.7} & {\it 3.1} \\ 
        $\mathcal{W}_{robot}$ & 89.6 & 84.4 & 86.6 & 72.8 & 81.9 & 79.7 & {\bf 82.5} & {\it 5.4} \\
        \bottomrule
    \end{tabular}
    \label{tab:results-with-me-ness}
\end{table}

Figure~\ref{fig:with-me-ness} can be further analyzed:\fixme{This paragraph
needs to be adapted}
The matching curve (cyan color) shows the overlapping between the
expected and the real focuses of attention. As can be noticed there is a
fluctuation at the beginning of the interaction as well as just before the
activity switching. These chaotic gaze manifestation is what we call an
automatic discovery of events, when the user experiences an uncertainty in the
present situation.
